Abstractions for Multi-Sorted Substitutions

### 2023-09-19 paper/paper.tex
Update.

  % \ExampleCode\FExampleTrav

  % \subsection{Unused}
  % \LibCode\KKitOpenInst
  % \LibCode\KKitOpen

@@

  Composition \AField{\_·ₖ\_} equips our framework with the power to
  reason about multiple maps \ACode{ϕ₁} and \ACode{ϕ₂} being applied
  to a term in sequence, in terms of a single application:

  The core property of composition is that applying two maps
  \ACode{ϕ₁} and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same
  as applying their composition \ACode{(ϕ₁ \AField{·ₖ} ϕ₂)} to \ACode{t}, i.e.
  $$
  \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
  $$

  Composition \AField{\_·ₖ\_} equips our framework with the power to
  reason about multiple maps \ACode{ϕ₁} and \ACode{ϕ₂} being applied
  to a term in sequence:

  \ACode{\AField{⋯-assoc} t ϕ₁ ϕ₂} \ \ : \ \ \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}

  For many lemmas, e.g. substitution-preserves-typing in System F,
  it is necessary to apply an equality like \ACode{t ⋯ σ₁ ⋯ σ₂ ≡ t ⋯ σ₂' ⋯ σ₁'}
  where the induction hypothesis requires \ACode{t' ⋯ σ₁'}.

  Composition of substitutions is important for most language
  formalizations, as it allows to reason about the meaning of applying
  multiple substitutions to a term. This appears for example in the proof of
  substitution-preserves-typing for System F, where in the
  case of type-application \ACode{(e \ACon{∙} t) \ADef{⋯} σ} we need to reason about
  \ACode{(e \ADef{⋯} \ADef{⦅} t \ADef{⦆}) \ADef{⋯} σ}, which is equivalent to
  \ACode{e \ADef{⋯} (\ADef{⦅} t \ADef{⦆} \ADef{·ₖ} σ)},
  where \ACode{\_\ADef{·ₖ}\_} is backward composition.

  As we defined substitution in terms of renamings, we need to
  consider all four compositions between renamings and substitutions.
  While the composition operations are defined independently of each other,
  the \ACode{\AField{⋯-assoc}} lemma for two substitutions, depends on the
  \ACode{\AField{⋯-assoc}} lemmas between a renaming and a substitution, which in
  turn depend on the \ACode{\AField{⋯-assoc}} lemma between two renamings.

  McBride et al.\cite{DBLP:journals/jar/BentonHKM12} addressed this
  issue by duplicating the definitions and using tactics to reduce
  boilerplate in proofs.

  In contrast we define structures similar to \ACode{\ACon{Kit}} and
  \ACode{\ACon{Traversal}}, which allows us to abstract over all four
  compositions and use the same trick as before to eliminate the
  dependencies.
  This has the benefit that a user of our library has to write less
  boilerplate, but more importantly it allows to prove lemmas about the
  composition of arbitrary maps, which enable us later to unify type
  preservation of all substitution and renaming operations into a single
  lemma.

@@

  \ACode{\ACon{ComposeKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
  \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-assoc}} are analogous to
  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.

@@

  \subsection{System F}
  \AppCode\FAssoc
  \AppCode\FAssocProofInteresting
  % \AppCode\FAssocProofRest
  \AppCode\FComposeTraversal

@@

  % \AppCode\FAssoc
  % \AppCode\FAssocProofInteresting
  % \AppCode\FAssocProofRest

>>>>

  % TODO note that its derived

  % \subsection{Unused}
  % \ExampleCode\FExampleTrav
  % \LibCode\KKitOpenInst
  % \LibCode\KKitOpen

@@

  The core property of composition is that applying two maps
  \ACode{ϕ₁} and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same
  as applying their composition \ACode{(ϕ₁ \AField{·ₖ} ϕ₂)} to \ACode{t}, i.e.
  $$
  \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
  $$
  We call this property \ACode{\AField{⋯-fusion}}.
  This property equips our framework with the power to
  reason about the application of multiple maps in terms of
  the application of a single map.
  As such it forms the basis for all lemmas involving multiple maps,
  e.g.\ that applying a weakening and then a singleton
  substitution cancel each other out.

  % For example, we can prove that first applying a weakening
  % \ACode{\ADef{weaken} s}, and then applying a singleton substitution
  % \ACode{\ADef{⦅} t \ADef{⦆}} cancel each other cancel each other out
  % by proving that
  % \ACode{(\ADef{weaken} s \AField{·ₖ} \ADef{⦅} t \ADef{⦆}) \ADef{$\sim$} id}.
  % As such it forms the basis for proving that a shifting cancels a
  % singleton substitution, or that

  % For many lemmas, e.g. substitution-preserves-typing in System F,
  % it is necessary to apply an equality like \ACode{t ⋯ σ₁ ⋯ σ₂ ≡ t ⋯ σ₂' ⋯ σ₁'}
  % where the induction hypothesis requires \ACode{t' ⋯ σ₁'}.

  % Composition of substitutions is important for most language
  % formalizations, as it allows to reason about the meaning of applying
  % multiple substitutions to a term. This appears for example in the proof of
  % substitution-preserves-typing for System F, where in the
  % case of type-application \ACode{(e \ACon{∙} t) \ADef{⋯} σ} we need to reason about
  % \ACode{(e \ADef{⋯} \ADef{⦅} t \ADef{⦆}) \ADef{⋯} σ}, which is equivalent to
  % \ACode{e \ADef{⋯} (\ADef{⦅} t \ADef{⦆} \ADef{·ₖ} σ)},
  % where \ACode{\_\ADef{·ₖ}\_} is backward composition.

  As we defined substitution in terms of renamings, we need to
  consider all four compositions between renamings and substitutions.
  While the composition operations can be defined independently of each other,
  the \ACode{\AField{⋯-fusion}} lemma for two substitutions, depends on the
  \ACode{\AField{⋯-fusion}} lemmas between a renaming and a substitution, which in
  turn depend on the \ACode{\AField{⋯-fusion}} lemma between two renamings.

  Benton et al.\cite{DBLP:journals/jar/BentonHKM12} addresses this
  issue by duplicating the definitions and using tactics to reduce
  boilerplate in proofs.
  In contrast, we define structures similar to \ACode{\ACon{Kit}} and
  \ACode{\ACon{Traversal}}, which allow us to abstract over all four
  compositions and use the same trick as before to eliminate the
  dependencies.
  This has the benefit, that we can prove lemmas about
  the composition of arbitrary maps, which enable us later to unify type
  preservation of all substitution and renaming operations into a single
  lemma.
  % TODO: clarify

@@

  \ACode{\ACon{ComposeKit}} and \ACode{\ADef{\_·ₖ\_}} are analogous to
  \ACode{\ACon{Kit}} and \ACode{\ADef{\_→ₖ\_}}, whereas
  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{⋯-fusion}} are analogous to
  \ACode{\ACon{ComposeTraversal}} and \ACode{\AField{\_⋯\_}}.

@@

  \subsection{System F}
  \AppCode\FFusion
  \AppCode\FFusionProofInteresting
  % \AppCode\FFusionProofRest
  \AppCode\FComposeTraversal

@@

  % \AppCode\FFusion
  % \AppCode\FFusionProofInteresting
  % \AppCode\FFusionProofRest
